---
title: "Acoustic Analysis"
author: "Xinyu Zhang"
output: html_notebook
---

```{r}
fulltable <- read.csv("results.csv")
library(dplyr)
nobins<-fulltable %>%
  select(Subject, Age, Index, IPA, Sibilant, CoG, Stdev,Position, Vp, Vf)
#changing the column "Sibilant" to logical:
nobins$Sibilant <- as.factor(nobins$Sibilant)
#changing the column "Age" to logical:
nobins$Age <- as.factor(nobins$Age)
#save the csv for appendix:
write.csv(nobins, file = "nobins.csv") #saving this because I might need it later for a different R script
nobins
```

```{r}
#checking whether the distribution is close enough to Gaussian
library(ggplot2)
qplot(hrxf$CoG, geom="histogram", bins = 150, xlab = "CoG (Hz)", ylab = "Count") 
qplot(hrxf$Stdev, geom="histogram", bins = 150, xlab = "Stdev (Hz)", ylab = "Count") 
```

```{r}
hrxf %>% 
 filter(heightP == "m")
```


```{r}
library(dplyr)
#categorizing preceding vowels:
nobins <- read.csv("nobins.csv")
hrxf <- nobins %>%mutate(
    heightP = case_when(
    Vp == "ɛin" ~"h",
    Vp == "y"~"h",
    Vp == "i" ~ "h",
    Vp == "u" ~"h",
    Vp == "ɛi" ~"h",
    Vp == "ɶy" ~"h",
    Vp == "au"~"h", #"a" w/ a hat but not distinguished because of font in R
    Vp == "ei"~"mh",
    Vp == "ou" ~"mh",
    Vp == "ə" ~ "m",
    Vp == "ɔ" ~ "ml",
    Vp == "ɛ" ~ "ml",
    Vp == "aɹ" ~ "l", #"a" w/ a hat but not distinguished because of font in R
    Vp == "a" ~ "l" #"a" w/ a hat but not distinguished because of font in R
    )
    )
hrxf$heightP <- as.factor(hrxf$heightP)

hrxf <- hrxf %>%mutate(
    heightF = case_when(
    Vf == "ɶy" ~"l",
    Vf == "au"~"l",
    Vf == "ɔ" ~ "ml",
    Vf == "a" ~ "l", #"a" w/ a hat but not distinguished because of font in R
    Vf == "ai" ~ "l", #"a" w/ a hat but not distinguished because of font in R
    Vf == "ɪ" ~ "mh",
    Vf == "o" ~ "mh",
    Vf == "i" ~ "h",
    Vf == "u" ~ "h",
    Vf == "ã" ~ "h",  #"a" w/ a hat but not distinguished because of font in R
    Vf == "y"~"h",
    Vf == "ə" ~ "m",
    Vf == "ɚ" ~ "m",
    Vf == "au" ~ "m" #"a" w/ a hat but not distinguished because of font in R
    )
    )

hrxf$heightF <- as.factor(hrxf$heightF)

hrxf <- hrxf %>%mutate(
roundingP = case_when(Vp == "ɔ" ~ "r",
                     Vp == "ou" ~ "r",
                     Vp == "ɶy" ~ "r",
                     Vp == "u" ~ "r",
                     Vp == "y" ~ "r",
                     Vp == "au" ~ "r", #"a" w/ a hat but not distinguished because of font in R
                     TRUE ~ "nr")
                     )

hrxf$roundingP <- as.factor(hrxf$roundingP)


hrxf <- hrxf %>%mutate(
roundingF = case_when(Vf == "ɔ" ~ "r",
                      Vf == "ɶy" ~ "r",
                      Vf == "o" ~ "r",
                      Vf == "u" ~ "r",
                      Vf == "y" ~ "r",
                      TRUE ~ "nr")
                      )

hrxf$roundingF <- as.factor(hrxf$roundingF)


hrxf <- hrxf %>%mutate(
rhoticityP = case_when( 
  Vp == "aɹ" ~ "x", #"a" w/ a hat but not distinguished because of font in R
  TRUE ~ "nx"
)
)

hrxf$rhoticityP <- as.factor(hrxf$rhoticityP)

hrxf <- hrxf %>%mutate(
rhoticityF = case_when( 
  Vf == "ɚ" ~ "x", 
  TRUE ~ "nx"
)
)

hrxf$rhoticityF <- as.factor(hrxf$rhoticityF)

#frontness of preceeding vowel
hrxf <- hrxf %>%mutate(
frtP = case_when( 
  Vf == "ɔ" ~ "b",
  Vf == "ou" ~ "b",
  Vf == "u" ~ "b",
  Vf == "au" ~ "b",
  Vf == "ə" ~ "m",
  TRUE ~ "f"
)
)

hrxf$frtP <- as.factor(hrxf$frtP)

#frontness of following vowel
hrxf <- hrxf %>%mutate(
frtF = case_when( 
  Vf == "ə" ~ "m",
  Vf == "ɔ" ~ "b",
  Vf == "o" ~ "b",
  Vf == "u" ~ "b",
  Vf == "ɚ" ~ "b",
  TRUE ~ "f"
)
)

hrxf$frtF <- as.factor(hrxf$frtF)


library(lme4)
#contrast coding for the height of the preceeding vowel, using the schwa as the reference level 
contrast1 <- cbind(c(-1/5, +4/5, -1/5,-1/5, -1/5),c(-1/5, -1/5, +4/5, -1/5,-1/5),c(-1/5, -1/5, -1/5, +4/5, -1/5), c(-1/5, -1/5,-1/5, -1/5, +4/5)) #m,h,mh,ml,l
colnames(contrast1) <- c("+h-mmhmll", "+mh-hmmll", "+ml-hmhml","+l-hmhmml")
contrasts(hrxf$heightP) <- contrast1

#contrast coding for the height of the following vowel, using the schwa as the reference level
contrast2 <- cbind(c(-1/5, +4/5, -1/5,-1/5, -1/5),c(-1/5, -1/5, +4/5, -1/5,-1/5),c(-1/5, -1/5, -1/5, +4/5, -1/5), c(-1/5, -1/5,-1/5, -1/5, +4/5)) #m,h,mh,ml,l
colnames(contrast2) <- c("+h-mmhmll", "+mh-hmmll", "+ml-hmhml","+l-hmhmml")
contrasts(hrxf$heightF) <- contrast2

#contrast coding for the rounding of the preceeding vowel
contrast3 <- cbind (c(-0.5, +0.5))
colnames (contrast3) <- c("-r+nr")
contrasts(hrxf$roundingP) <- contrast3

#contrast coding for the rounding of the following vowel
contrast4 <- cbind (c(-0.5, +0.5))
colnames (contrast4) <- c("-r+nr")
contrasts(hrxf$roundingF) <- contrast4

#contrast coding for the rhoticity of the preceeding vowel
contrast5 <- cbind (c(-0.5, +0.5))
colnames (contrast5) <- c("-x+nx")
contrasts(hrxf$rhoticityP) <- contrast5

#contrast coding for the rhoticity of the following vowel
contrast6 <- cbind (c(-0.5, +0.5))
colnames (contrast6) <- c("-x+nx")
contrasts(hrxf$rhoticityF) <- contrast6

#contrast coding for the frontness of the preceeding vowel, using the schwa as the reference level
contrast7 <- cbind (c(-1/3, +2/3, -1/3), c(-1/3, -1/3, +2/3))#m,f,b
colnames (contrast7) <- c("+f-mb", "+b,-mf")
contrasts(hrxf$frtP) <- contrast7

#contrast coding for the frontness of the following vowel, using the schwa as the reference level
contrast8 <- cbind (c(-1/3, +2/3, -1/3), c(-1/3, -1/3, +2/3))#m,f,b
colnames (contrast8) <- c("+f-mb", "+b,-mf")
contrasts(hrxf$frtF) <- contrast8

#contrast coding for position, using the medials as the reference level 
hrxf$Position <- as.factor(hrxf$Position)
contrast9 <- cbind (c(-1/3, +2/3, -1/3), c(-1/3, -1/3, +2/3))#m,i,f
colnames (contrast9) <- c("+i-mf", "+f,-im")
contrasts(hrxf$Position) <- contrast9

#model
#trying with only the initials first, to make the model more likely to be runnable
initials <- hrxf %>% 
  filter(Position == "initial")

fullModel <- glmer(Age ~ Sibilant*CoG*Stdev*heightP*heightF*roundingP*roundingF*rhoticityP*rhoticityF*frtP*frtF + (CoG*Stdev*heightP*heightF*roundingP*roundingF*rhoticityP*rhoticityF*frtP*frtF|Subject) + (Sibilant*CoG*Stdev*heightP*heightF*roundingP*roundingF*rhoticityP*rhoticityF*frtP*frtF|Index), data = initials, family = binomial)
summary(fullModel)
```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

